---
title: "Balto-Slavic"
editor: visual
---

```{r}
#| echo: false
#| messages: false
#| output: false

library(tidyverse)
library(mstknnclust)
library(corrplot)
data("dslanguages")
```

```{r}
#| echo: false

##First I grouped the languages into a branch to make intra-branch comparisons possible
balto_slavic <- dslanguages |>
  select(Latvian:Serbocroatian)

##This sets the data up for the matrix and then gives the languages more modern names (the data is from 1992)
balto_slavic_plot <- cor(balto_slavic)

colnames(balto_slavic_plot) <- rownames(balto_slavic_plot) <- c(
                                "Latvian", "Slovenian", 
                                "Upper Sorbian", "Lower Sorbian", 
                                "Czech", "Slovak", 
                                "East Czech (Trnovec dialect)",
                                "Ukrainian", "Belarusian", "Polish",
                                "Russian", "Macedonian", "Bulgarian",
                                "Serbo-Croatian")
####This actually creates to the plot.  is.corr MUST equal FALSE to get the coloring right.
corrplot(balto_slavic_plot, method = 'color', is.corr = FALSE)
```

This is an internal comparison of the Balto-Slavic branch, the most politically fraught of the bunch. "Serbo-Croatian" represents all of the dialects spoken in the former Yugoslavia, which are mutually intelligible (the usual test of whether two dialects form a single language) but which the region's governments insist are actually separate Serbian, Bosnian, Croatian, and Montenegrin languages. Some Lithuanians and Latvians may resent being grouped in with the Slavic languages, associated particularly with Pan-Slavist heimat and former occupier Russia, but the linguistic case is closed here. Note that Polish and Slovene are just as Slavic as Russian is.

The languages are all very similar to each other as you can see on the bar to the right of the chart, but Latvian stands out as the most distinct. This is almost certainly because Balto-Slavic is divided into two sub-branches, Baltic and Slavic, with Latvian the only representative of the former in this dataset. Other than this the range is so restricted that I don't think the data can provide many insights.
