---
title: "Lexical Similarities Between Different Branches of Indo-European"
---
```{r}
#| echo: false
#| messages: false
#| output: false

library(tidyverse)
library(mstknnclust)
library(corrplot)
data("dslanguages")
```


```{r}
#| echo: false
#| messages: false

##This code is found on other docs, but it renders more efficiently when it's laid out here.

##First I grouped the languages according to branches, as my goal is primarily to compare the different branches.

celtic <- dslanguages |>
  select(IrishA, IrishB, WelshN, WelshC, 
         BretonList, BretonSE, BretonST)

romance <- dslanguages |>
  select(RumanianList:Catalan)


germanic <- dslanguages |>
  select(GermanST:Takitaki)


balto_slavic <- dslanguages |>
  select(Latvian:Serbocroatian)

indic <- dslanguages |>
select(GypsyGk:Khaskura)


greek <- dslanguages |>
  select(GreekML:GreekK)

armenian <- dslanguages |>
  select(ArmenianMod:ArmenianList)

iranic <- dslanguages |>
  select(Ossetic:Wakhi)

albanian <- dslanguages |>
  select(AlbanianT:AlbanianC)

##Then I created a "branch" variable.  select() didn't work, so I used list()

branch <- list(celtic, romance, germanic, balto_slavic, indic,
               greek, armenian, iranic, albanian)

##These are necessary to get the various branches (not the branch variable!) to have numerical values we can compare.

row_sums <- lapply(branch, rowSums)
row_sums_matrix <- do.call(cbind, row_sums)

##The next two lines set up the correlational variable and give the colums appropriate names.
M <- cor(row_sums_matrix)
colnames(M) <- rownames(M) <- c("Celtic", "Romance", 
                                "Germanic", "Balto-Slavic", "Indic",
                                "Greek", "Armenian", "Iranic", "Albanian")

##This actually creates to the plot.  is.corr MUST equal FALSE to get the coloring right.
corrplot(M, method = 'color', is.corr = FALSE)


```

The Indo-European language family is the largest group of languages on the planet, with English, French, Russian, Greek and Hindi all included in its ranks. I used Dyen et al.'s (1992) dataset of lexical distances between 82 Indo-European languages to investigate how similar the vocabularies of the different branches (major groups) of languages are.  The chart above groups those languages according to genetic branches and calculates the average similarities between the different languages of these larger groups; linguists usually care about how close two given langauges are to a common ancestor (their genetic relatedness), but my project only compares how similar their common words are. European branches are generally more similar to each other than they are to Asian branches, despite being no more closely related.