---
title: "Lexical Similarities Between Different Branches of Indo-European"
---
```{r}
#| echo: false
#| messages: false
#| output: false

library(tidyverse)
library(mstknnclust)
library(corrplot)
data("dslanguages")
```


```{r}
#| echo: false
#| messages: false

##This code is found on other docs, but it renders more efficiently when it's laid out here.

##First I grouped the languages according to branches, as my goal is primarily to compare the different branches.

celtic <- dslanguages |>
  select(IrishA, IrishB, WelshN, WelshC, 
         BretonList, BretonSE, BretonST)

romance <- dslanguages |>
  select(RumanianList:Catalan)


germanic <- dslanguages |>
  select(GermanST:Takitaki)


balto_slavic <- dslanguages |>
  select(Latvian:Serbocroatian)

indic <- dslanguages |>
select(GypsyGk:Khaskura)


greek <- dslanguages |>
  select(GreekML:GreekK)

armenian <- dslanguages |>
  select(ArmenianMod:ArmenianList)

iranic <- dslanguages |>
  select(Ossetic:Wakhi)

albanian <- dslanguages |>
  select(AlbanianT:AlbanianC)

##Then I created a "branch" variable.  select() didn't work, so I used list()

branch <- list(celtic, romance, germanic, balto_slavic, indic,
               greek, armenian, iranic, albanian)

##These are necessary to get the various branches (not the branch variable!) to have numerical values we can compare.

row_sums <- lapply(branch, rowSums)
row_sums_matrix <- do.call(cbind, row_sums)

##The next two lines set up the correlational variable and give the colums appropriate names.
M <- cor(row_sums_matrix)
colnames(M) <- rownames(M) <- c("Celtic", "Romance", 
                                "Germanic", "Balto-Slavic", "Indic",
                                "Greek", "Armenian", "Iranic", "Albanian")

##This actually creates to the plot.  is.corr MUST equal FALSE to get the coloring right.
corrplot(M, method = 'color', is.corr = FALSE)


```

This chart shows how similar the vocabularies of different branches of the Indo-European language family are. The data is taken from fron Dyen et al.'s (1992) matrix of the lexical distances between Indo-European languages.  My chart groups those languages according to genetic branches and calculates the average similarities between the different languages of these larger groups. European branches are generally more similar to each other than they are to Asian branches, despite being no more closely related.